---
title: "pricing bond options"
author: "mdowns"
date: "November 04, 2015"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='hide',warning=FALSE,fig.show='asis',fig.height=3.5,fig.width=5}
library(xts)
library(urca)
library(quantmod)
library(fUnitRoots)
library(PerformanceAnalytics)
library(highfrequency)
library(fOptions)
library(optimx)
#library(timeSeries)
#library(timeDate)

#library(lubridate)
#library(tseries)
#library(TTR)
#library(caret)
#library(mondate)
#library(MTS)
#library(car)

```

## problem 1

*(a) Use these data to calibrate the model parameters ($\kappa$, $\theta$, $\sigma$) and instantaneous interest rates on December 20, 2006.*

Parameter values are below. The first set is from the unconstrained optimization. The second set is from the constrained optimization.

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

# retrieve ddata
address.head <- "http://www.stanford.edu/~xing/statfinbook/_BookData/"
address.tail <- "Chap10/bonds_yield_dec2006.txt"
address <- paste(address.head,address.tail,sep='')
data <- read.table(address,head=T)

# functions
est_beta_vs=function(kappa,ttm){(1-exp(-(kappa*ttm)))/kappa}
est_alpha_vs=function(theta,sigma,kappa,beta,ttm){
     exp((theta-(sigma^2/(2*(kappa^2))))*(beta-ttm)-((sigma^2)/(4*kappa))*beta^2)}
est_r_t=function(alpha,beta,ttm,r_t){-(log(alpha)/ttm)+r_t*(beta/ttm)}

# start values
r_t=data$yield/100
kappa=0.5
theta=mean(r_t)
sigma=mean(r_t)
ttm=data$year.to.maturity

parms=as.data.frame(t(c(kappa,theta,sigma)));colnames(parms)=c("kappa","theta","sigma")
dat=as.data.frame(cbind(ttm,r_t))

# lease squares fit function
opt_mdl_vs=function(data,par){
     
     kappa=par[1]
     theta=par[2]
     sigma=par[3]
     ttm=data[,1]
     r_t=data[,2]
     
     beta=est_beta_vs(kappa,ttm)
     alpha=est_alpha_vs(theta,sigma,kappa,beta,ttm)
     spot=est_r_t(alpha,beta,ttm,r_t)
     
     fit.err=sum((r_t-spot)^2)
     return(fit.err)
}

# unconstrained optimization
rslt.vs.opt.unc=optim(par=parms,fn=opt_mdl_vs,data=dat)
rslt.vs.opt.unc$par
kappa.vs.opt.unc=rslt.vs.opt.unc$par[1];theta.vs.opt.unc=rslt.vs.opt.unc$par[2];sigma.vs.opt.unc=rslt.vs.opt.unc$par[3]

# constrained optimization
rslt.vs.opt.c=optim(par=parms,fn=opt_mdl_vs,data=dat,lower=c(0.0001,0.0001,0.0001))
rslt.vs.opt.c$par
kappa.vs.opt.c=rslt.vs.opt.c$par[1];theta.vs.opt.c=rslt.vs.opt.c$par[2];sigma.vs.opt.c=rslt.vs.opt.c$par[3]

```

Both instantaneous and spot rates are shown in the graphics below. 

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=3.5,fig.width=7}
#******** rates
# spot rates
beta.vs.opt.c=est_beta_vs(kappa.vs.opt.c,ttm)
alpha.vs.opt.c=est_alpha_vs(theta.vs.opt.c,sigma.vs.opt.c,kappa.vs.opt.c,beta.vs.opt.c,ttm)
spot.vs.opt.c=est_r_t(alpha.vs.opt.c,beta.vs.opt.c,ttm,r_t)

beta.vs.opt.unc=est_beta_vs(kappa.vs.opt.unc,ttm)
alpha.vs.opt.unc=est_alpha_vs(theta.vs.opt.unc,sigma.vs.opt.unc,kappa.vs.opt.unc,beta.vs.opt.unc,ttm)
spot.vs.opt.unc=est_r_t(alpha.vs.opt.unc,beta.vs.opt.unc,ttm,r_t)

# instantaneous rates
library(numDeriv)

## unconstrained kappa, sigma, theta provided by optim
beta_func_vs_unc=function(ttm){(1-exp(-(-0.000129566*ttm)))/-0.000129566}

alpha_func_inst_vs_unc=function(ttm){log(exp((0.0470205450-(0.0001627793^2/(2*(-0.000129566^2))))*(beta_func_vs_unc(ttm)-ttm)-
                                         ((0.0001627793^2)/(4*-0.000129566))*beta_func_vs_unc(ttm)^2))}

alpha_func_vs_unc=function(ttm){exp((0.04717334-(0.0001627793^2/(2*(-0.000129566^2))))*(beta_func_vs_unc(ttm)-ttm)-
                                         ((0.0001627793^2)/(4*-0.000129566))*beta_func_vs_unc(ttm)^2)}

## constrained kappa, sigma, theta provided by optim
beta_func_vs_c=function(ttm){(1-exp(-(0.50156643*ttm)))/0.50156643}

alpha_func_inst_vs_c=function(ttm){log(exp((0.04690427-(0.00010000^2/(2*(0.50156643^2))))*(beta_func_vs_c(ttm)-ttm)-
                                         ((0.00010000^2)/(4*0.50156643))*beta_func_vs_c(ttm)^2))}

alpha_func_vs_c=function(ttm){exp((0.04690427-(0.00010000^2/(2*(0.50156643^2))))*(beta_func_vs_c(ttm)-ttm)-
                                         ((0.00010000^2)/(4*0.50156643))*beta_func_vs_c(ttm)^2)}

## calculate the gradient
gb.c=grad(func=beta_func_vs_c,x=ttm)
ga.c=grad(func=alpha_func_inst_vs_c,x=ttm)
rslt.c=-ga.c + r_t*gb.c

gb.unc=grad(func=beta_func_vs_unc,x=ttm)
ga.unc=grad(func=alpha_func_inst_vs_unc,x=ttm)
rslt.unc=-ga.unc + r_t*gb.unc

## plot the results
par(mfrow=c(1,2))
plot(r_t,type="l",lwd=2,main="constrained vasicek yields",ylim=range(cbind(spot.vs.opt.c,r_t)))
lines(spot.vs.opt.c,col="red",lwd=2)
lines(rslt.c,col="blue",lwd=2)
legend("topright",legend=c("observed","spot - optim","instant - optim"),col=c("black","red","blue"),lwd=2,cex=0.7)

plot(r_t,type="l",lwd=2,main="unconstrained vasicek yields",ylim=range(cbind(spot.vs.opt.unc,rslt.unc)))
lines(spot.vs.opt.unc,col="red",lwd=2)
lines(rslt.unc,col="blue",lwd=2)
legend("topright",legend=c("observed","spot - optim","instant - optim"),col=c("black","red","blue"),lwd=2,cex=0.7)

```

\textbf{comments:} 

The key to getting any result in part "b" below was to constrain the least squares optimization of $\kappa$, $\sigma$ and $\theta$ to positive numbers. So, the graphics above show the computation of spot and instantaneous rates using constrained and unconstrained models. Note that the observed rates in the unconstrained Vasicek yields have been overwritten by the optimized Vasicek fit (red) which interpolated the data. Error rates for the unconstrained and constrained fits, respectively, are:

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

rslt.vs.opt.unc$value;rslt.vs.opt.c$value

```

\textbf{code:}
```{r eval=FALSE,cache=FALSE,echo=TRUE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

# retrieve ddata
address.head <- "http://www.stanford.edu/~xing/statfinbook/_BookData/"
address.tail <- "Chap10/bonds_yield_dec2006.txt"
address <- paste(address.head,address.tail,sep='')
data <- read.table(address,head=T)

# functions
est_beta_vs=function(kappa,ttm){(1-exp(-(kappa*ttm)))/kappa}
est_alpha_vs=function(theta,sigma,kappa,beta,ttm){
     exp((theta-(sigma^2/(2*(kappa^2))))*(beta-ttm)-((sigma^2)/(4*kappa))*beta^2)}
est_r_t=function(alpha,beta,ttm,r_t){-(log(alpha)/ttm)+r_t*(beta/ttm)}

# start values
r_t=data$yield/100
kappa=0.5
theta=mean(r_t)
sigma=mean(r_t)
ttm=data$year.to.maturity

parms=as.data.frame(t(c(kappa,theta,sigma)));colnames(parms)=c("kappa","theta","sigma")
dat=as.data.frame(cbind(ttm,r_t))

# lease squares fit function
opt_mdl_vs=function(data,par){
     
     kappa=par[1]
     theta=par[2]
     sigma=par[3]
     ttm=data[,1]
     r_t=data[,2]
     
     beta=est_beta_vs(kappa,ttm)
     alpha=est_alpha_vs(theta,sigma,kappa,beta,ttm)
     spot=est_r_t(alpha,beta,ttm,r_t)
     
     fit.err=sum((r_t-spot)^2)
     return(fit.err)
}

# unconstrained optimization
rslt.vs.opt.unc=optim(par=parms,fn=opt_mdl_vs,data=dat)
rslt.vs.opt.unc$par
kappa.vs.opt.unc=rslt.vs.opt.unc$par[1];theta.vs.opt.unc=rslt.vs.opt.unc$par[2];sigma.vs.opt.unc=rslt.vs.opt.unc$par[3]

# constrained optimization
rslt.vs.opt.c=optim(par=parms,fn=opt_mdl_vs,data=dat,lower=c(0.0001,0.0001,0.0001))
rslt.vs.opt.c$par
kappa.vs.opt.c=rslt.vs.opt.c$par[1];theta.vs.opt.c=rslt.vs.opt.c$par[2];sigma.vs.opt.c=rslt.vs.opt.c$par[3]

#******** rates
# spot rates
beta.vs.opt.c=est_beta_vs(kappa.vs.opt.c,ttm)
alpha.vs.opt.c=est_alpha_vs(theta.vs.opt.c,sigma.vs.opt.c,kappa.vs.opt.c,beta.vs.opt.c,ttm)
spot.vs.opt.c=est_r_t(alpha.vs.opt.c,beta.vs.opt.c,ttm,r_t)

beta.vs.opt.unc=est_beta_vs(kappa.vs.opt.unc,ttm)
alpha.vs.opt.unc=est_alpha_vs(theta.vs.opt.unc,sigma.vs.opt.unc,kappa.vs.opt.unc,beta.vs.opt.unc,ttm)
spot.vs.opt.unc=est_r_t(alpha.vs.opt.unc,beta.vs.opt.unc,ttm,r_t)

# instantaneous rates
library(numDeriv)

## unconstrained kappa, sigma, theta provided by optim
beta_func_vs_unc=function(ttm){(1-exp(-(-0.000129566*ttm)))/-0.000129566}

alpha_func_inst_vs_unc=function(ttm){log(exp((0.0470205450-(0.0001627793^2/(2*(-0.000129566^2))))*(beta_func_vs_unc(ttm)-ttm)-
                                         ((0.0001627793^2)/(4*-0.000129566))*beta_func_vs_unc(ttm)^2))}

alpha_func_vs_unc=function(ttm){exp((0.04717334-(0.0001627793^2/(2*(-0.000129566^2))))*(beta_func_vs_unc(ttm)-ttm)-
                                         ((0.0001627793^2)/(4*-0.000129566))*beta_func_vs_unc(ttm)^2)}

## constrained kappa, sigma, theta provided by optim
beta_func_vs_c=function(ttm){(1-exp(-(0.50156643*ttm)))/0.50156643}

alpha_func_inst_vs_c=function(ttm){log(exp((0.04690427-(0.00010000^2/(2*(0.50156643^2))))*(beta_func_vs(ttm)-ttm)-
                                         ((0.00010000^2)/(4*0.50156643))*beta_func_vs_c(ttm)^2))}

alpha_func_vs_c=function(ttm){exp((0.04690427-(0.00010000^2/(2*(0.50156643^2))))*(beta_func_vs(ttm)-ttm)-
                                         ((0.00010000^2)/(4*0.50156643))*beta_func_vs_c(ttm)^2)}

## calculate the gradient
gb.c=grad(func=beta_func_vs_c,x=ttm)
ga.c=grad(func=alpha_func_inst_vs_c,x=ttm)
rslt.c=-ga.c + r_t*gb.c

gb.unc=grad(func=beta_func_vs_unc,x=ttm)
ga.unc=grad(func=alpha_func_inst_vs_unc,x=ttm)
rslt.unc=-ga.unc + r_t*gb.unc

## plot the results
par(mfrow=c(1,2))
plot(r_t,type="l",lwd=2,main="constrained vasicek yields",ylim=range(cbind(spot.vs.opt.c,r_t)))
lines(spot.vs.opt.c,col="red",lwd=2)
lines(rslt.c,col="blue",lwd=2)
legend("topright",legend=c("observed","spot - optim","instant - optim"),col=c("black","red","blue"),lwd=2)

plot(r_t,type="l",lwd=2,main="unconstrained vasicek yields",ylim=range(cbind(spot.vs.opt.unc,rslt.unc)))
lines(spot.vs.opt.unc,col="red",lwd=2)
lines(rslt.unc,col="blue",lwd=2)
legend("topright",legend=c("observed","spot - optim","instant - optim"),col=c("black","red","blue"),lwd=2)

```

*(b) With the fitted model and the initial short-term interest rate 4.75%, calculate the price of a 2.5-year European call option, with strike price $99, on a bond that matures in three years, pays coupons semiannually at 5% rate and has par value $100.*

Value of the 2.5 year european call option in dollars:

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

### given
bnd.r.t=0.0475 # short term rate
bnd.par=1 # par or face value of bond
bnd.ttm=3
div=0.05 
omega=1 # call option indicator
opt.strike=99/100 # strike price
opt.ttm=2.5 # option ttm

#********** bond price
price_func_vs=function(ttm){alpha_func_vs_c(ttm) * exp(-beta_func_vs_c(ttm)*r_t)}
p_t_vs=price_func_vs(ttm)*bnd.par

#********** option price
# vs-specific settings
bnd.beta.T.Tilde=(beta_func_vs_c(opt.ttm)-beta_func_vs_c(bnd.ttm)) # beta on 0.5 yr bond 2.5 yrs hence
bnd.p.t.Tilde=p_t_vs[which(ttm>3 & ttm<3.5)] #price today for bond maturing in three years
bnd.p.t.T=p_t_vs[which(ttm>2.5 & ttm<3)] # price today for bond maturing in 2.5 years

# Jamshidian bond option prices
sigma.p.vs=sigma.vs.opt.c*sqrt((1-exp(-2*kappa.vs.opt.c*opt.ttm))/(2*kappa.vs.opt.c))*bnd.beta.T.Tilde
#v2=sigma.vs.opt.c^2*(1-exp(-2*kappa.vs.opt.c*bnd.ttm))/2*kappa.vs.opt.c 
#sigma.p.vs=sqrt(v2)*(1-exp(-kappa.vs.opt.c*bnd.ttm))/kappa.vs.opt.c
h.vs=(log(bnd.p.t.Tilde/(bnd.p.t.T*opt.strike))/sigma.p.vs+sigma.p.vs/2)

z.t.vs=omega*((bnd.p.t.Tilde*pnorm(omega*h.vs))-(opt.strike*bnd.p.t.T*pnorm(omega*(h.vs-sigma.p.vs))))
as.numeric(z.t.vs*100)

```

\textbf{comments:}

The graphic below shows the expected bond values underlying the call option.

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=3.5,fig.width=5}

plot(ttm,p_t_vs,type="l",lwd=2,main="vasicek zero coupon bond prices")

```

\textbf{code:}
```{r eval=FALSE,cache=FALSE,echo=TRUE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

### given
bnd.r.t=0.0475 # short term rate
bnd.par=1 # par or face value of bond
bnd.ttm=3
div=0.05 
omega=1 # call option indicator
opt.strike=99/100 # strike price
opt.ttm=2.5 # option ttm

#********** bond price
price_func_vs=function(ttm){alpha_func_vs_c(ttm) * exp(-beta_func_vs_c(ttm)*r_t)}
p_t_vs=price_func_vs(ttm)*bnd.par

#********** option price
# vs-specific settings
bnd.beta.T.Tilde=(beta_func_vs_c(opt.ttm)-beta_func_vs_c(bnd.ttm)) # beta on 0.5 yr bond 2.5 yrs hence
bnd.p.t.Tilde=p_t_vs[which(ttm>3 & ttm<3.5)] #price today for bond maturing in three years
bnd.p.t.T=p_t_vs[which(ttm>2.5 & ttm<3)] # price today for bond maturing in 2.5 years

# Jamshidian bond option prices
sigma.p.vs=sigma.vs.opt.c*sqrt((1-exp(-2*kappa.vs.opt.c*opt.ttm))/(2*kappa.vs.opt.c))*bnd.beta.T.Tilde
#v2=sigma.vs.opt.c^2*(1-exp(-2*kappa.vs.opt.c*bnd.ttm))/2*kappa.vs.opt.c 
#sigma.p.vs=sqrt(v2)*(1-exp(-kappa.vs.opt.c*bnd.ttm))/kappa.vs.opt.c
h.vs=(log(bnd.p.t.Tilde/(bnd.p.t.T*opt.strike))/sigma.p.vs+sigma.p.vs/2)

z.t.vs=omega*((bnd.p.t.Tilde*pnorm(omega*h.vs))-(opt.strike*bnd.p.t.T*pnorm(omega*(h.vs-sigma.p.vs))))
as.numeric(z.t.vs*100)

```


*(c) Consider the same problem as in (a) and (b), but use the CIR model.*

Parameter values are below. The first set is from the unconstrained optimization. The second set is from the constrained optimization.

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=3.5,fig.width=5}

#********* 
# start values
r_t=data$yield/100
kappa=0.5
theta=mean(r_t)
sigma=sd(r_t)
ttm=data$year.to.maturity

parms=as.data.frame(cbind(kappa,theta,sigma));colnames(parms)=c("kappa","theta","sigma")
dat=as.data.frame(cbind(ttm,r_t))

# functions
est_h_cir=function(kappa,sigma){sqrt(kappa^2+2*(sigma^2))}

est_beta_cir=function(h,kappa,ttm){
     (2*(exp(ttm*h)-1))/
          (2*h+(kappa+h)*(exp(ttm*h)-1))}

est_alpha.cir=function(h,kappa,theta,sigma,ttm){
     ((2*h*exp(((kappa+h)*ttm)/2))/
           (2*h+(kappa+h)*(exp(ttm*h)-1)))^(2*kappa*(theta/sigma^2))}

# lease squares fit function
opt_mdl_cir=function(data,par){
     
     kappa=par[1]
     theta=par[2]
     sigma=par[3]
     ttm=data[,1]
     r_t=data[,2]
     
     h=est_h_cir(kappa,sigma)
     beta=est_beta_cir(h,kappa,ttm)
     alpha=est_alpha.cir(h,kappa,theta,sigma,ttm)
     spot=est_r_t(alpha,beta,ttm,r_t)
     
     fit.err=sum((r_t-spot)^2)
     return(fit.err)
}

# unconstrained optimization
rslt.cir.opt.u=optim(par=parms,fn=opt_mdl_cir,data=dat)
rslt.cir.opt.u$par
kappa.cir.opt.u=rslt.cir.opt.u$par[1];theta.cir.opt.u=rslt.cir.opt.u$par[2];sigma.cir.opt.u=rslt.cir.opt.u$par[3]

# constrained optimization
rslt.cir.opt=optim(par=parms,fn=opt_mdl_cir,data=dat,lower=c(0.0001,0.0001,0.0001))
rslt.cir.opt$par
kappa.cir.opt=rslt.cir.opt$par[1];theta.cir.opt=rslt.cir.opt$par[2];sigma.cir.opt=rslt.cir.opt$par[3]

```

Both instantaneous and spot rates are shown in the graphic below.

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=3.5,fig.width=7}

#******** rates
h.cir.opt.u=est_h_cir(kappa.cir.opt.u,sigma.cir.opt.u)
beta.cir.opt.u=est_beta_cir(h.cir.opt.u,kappa.cir.opt.u,ttm)
alpha.cir.opt.u=est_alpha.cir(h.cir.opt.u,kappa.cir.opt.u,theta.cir.opt.u,sigma.cir.opt.u,ttm)
spot.cir.opt.u=est_r_t(alpha.cir.opt.u,beta.cir.opt.u,ttm,r_t)

h.cir.opt=est_h_cir(kappa.cir.opt,sigma.cir.opt)
beta.cir.opt=est_beta_cir(h.cir.opt,kappa.cir.opt,ttm)
alpha.cir.opt=est_alpha.cir(h.cir.opt,kappa.cir.opt,theta.cir.opt,sigma.cir.opt,ttm)
spot.cir.opt=est_r_t(alpha.cir.opt,beta.cir.opt,ttm,r_t)

## instantaneous rates
beta_func_cir_u=function(ttm){(1-exp(-(-0.0000156*ttm)))/-0.0000156}

alpha_func_cir_u=function(ttm){exp((0.0478709-(0.000035175^2/(2*(-0.0000156^2))))*(beta_func_cir_u(ttm)-ttm)-
                                          (0.000035175^2/(4*-0.0000156))*beta_func_cir_u(ttm)^2)}

beta_func_cir=function(ttm){(1-exp(-(0.44977704*ttm)))/0.44977704}

alpha_func_cir=function(ttm){exp((0.04692863-(0.0001000^2/(2*(0.44977704^2))))*(beta_func_cir(ttm)-ttm)-
                                          (0.0001000^2/(4*0.44977704))*beta_func_cir(ttm)^2)}

## calculate the gradient
gb.u=grad(func=beta_func_cir_u,x=ttm,)
ga.u=grad(func=alpha_func_cir_u,x=ttm)
rslt.u=-ga.u + r_t*gb.u

gb=grad(func=beta_func_cir,x=ttm,)
ga=grad(func=alpha_func_cir,x=ttm)
rslt=-ga + r_t*gb

par(mfrow=c(1,2))
plot(r_t,type="l",lwd=2,main="constrained cir yields",ylim=range(cbind(spot.cir.opt,rslt)))
lines(spot.cir.opt,col="red",lwd=2)
lines(rslt,col="blue",lwd=2)
legend("bottomleft",legend=c("observed","spot - optim","instant - opim"),col=c("black","red","blue"),lwd=2,cex=0.7)

plot(r_t,type="l",lwd=2,main="unconstrained cir yields",ylim=range(cbind(spot.cir.opt.u,rslt.u)))
lines(spot.cir.opt.u,col="red",lwd=2)
lines(rslt.u,col="blue",lwd=2)
legend("topright",legend=c("observed","spot - optim","instant - opim"),col=c("black","red","blue"),lwd=2,cex=0.7)

```

Estimated value for the call option using cir:

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

#********** bond price
price_func_cir=function(ttm){alpha_func_cir(ttm) * exp(-beta_func_cir(ttm)*r_t)}
p_t_cir=price_func_cir(ttm)

#********** option price
bnd.beta.T.Tilde=beta_func_cir(opt.ttm)-beta_func_cir(bnd.ttm) # beta on 0.5 yr bond 2.5 yrs hence
bnd.alpha.T.Tilde=alpha_func_cir(opt.ttm)-alpha_func_cir(bnd.ttm)
bnd.p.t.Tilde=p_t_cir[which(ttm>3 & ttm<3.5)] #p_t_cir where 3 < ytm < 3.5
bnd.p.t.T=p_t_cir[which(ttm>2.5 & ttm<3)]

# bond option prices using cir
cir.rho=(2*h.cir.opt)/
     (sigma.cir.opt^2*(exp(h.cir.opt*opt.ttm)-1))

cir.psi=(kappa.cir.opt+h.cir.opt)/sigma.cir.opt^2

cir.mu=(log(bnd.alpha.T.Tilde/opt.strike))/
     bnd.beta.T.Tilde

cir.z=bnd.p.t.Tilde*pchisq(2*cir.mu*(cir.rho+cir.psi+bnd.beta.T.Tilde),
                           (4*kappa.cir.opt*theta.cir.opt)/sigma.cir.opt^2,
                           (2*cir.rho^2*bnd.r.t*exp(h.cir.opt*opt.ttm))/(cir.rho+cir.psi+bnd.beta.T.Tilde))-
     opt.strike*bnd.p.t.T*pchisq(2*cir.mu*(cir.rho+cir.psi),
                                 (4*kappa.cir.opt*theta.cir.opt)/sigma.cir.opt^2,
                                 (2*cir.rho^2*bnd.r.t*exp(h.cir.opt*opt.ttm))/(cir.rho+cir.psi))

as.numeric(cir.z*100)

```

\textbf{comments:}

The call value is based on the cir (black) bond prices below. Vasicek values (red) provided for reference. 

```{r eval=TRUE,cache=TRUE,echo=FALSE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=3.5,fig.width=5}

plot(p_t_cir,type="l",main="bond prices: cir v vasicek",lwd=2,ylab="presumably this is discount to face value")
lines(p_t_vs,col="red",lwd=2)
legend("bottomleft",legend=c("cir - optim","vasicek - optim"),col=c("black","red"),lwd=2)

```

\textbf{code:}
```{r eval=FALSE,cache=FALSE,echo=TRUE,message=FALSE,results='markup',warning=FALSE,fig.show='asis',fig.height=4,fig.width=7}

#********* 
# start values
r_t=data$yield/100
kappa=0.5
theta=mean(r_t)
sigma=sd(r_t)
ttm=data$year.to.maturity

parms=as.data.frame(cbind(kappa,theta,sigma));colnames(parms)=c("kappa","theta","sigma")
dat=as.data.frame(cbind(ttm,r_t))

# functions
est_h_cir=function(kappa,sigma){sqrt(kappa^2+2*(sigma^2))}

est_beta_cir=function(h,kappa,ttm){
     (2*(exp(ttm*h)-1))/
          (2*h+(kappa+h)*(exp(ttm*h)-1))}

est_alpha.cir=function(h,kappa,theta,sigma,ttm){
     ((2*h*exp(((kappa+h)*ttm)/2))/
           (2*h+(kappa+h)*(exp(ttm*h)-1)))^(2*kappa*(theta/sigma^2))}

# lease squares fit function
opt_mdl_cir=function(data,par){
     
     kappa=par[1]
     theta=par[2]
     sigma=par[3]
     ttm=data[,1]
     r_t=data[,2]
     
     h=est_h_cir(kappa,sigma)
     beta=est_beta_cir(h,kappa,ttm)
     alpha=est_alpha.cir(h,kappa,theta,sigma,ttm)
     spot=est_r_t(alpha,beta,ttm,r_t)
     
     fit.err=sum((r_t-spot)^2)
     return(fit.err)
}

# unconstrained optimization
rslt.cir.opt.u=optim(par=parms,fn=opt_mdl_cir,data=dat)
rslt.cir.opt.u$par
kappa.cir.opt.u=rslt.cir.opt.u$par[1];theta.cir.opt.u=rslt.cir.opt.u$par[2];sigma.cir.opt.u=rslt.cir.opt.u$par[3]

# constrained optimization
rslt.cir.opt=optim(par=parms,fn=opt_mdl_cir,data=dat,lower=c(0.0001,0.0001,0.0001))
rslt.cir.opt$par
kappa.cir.opt=rslt.cir.opt$par[1];theta.cir.opt=rslt.cir.opt$par[2];sigma.cir.opt=rslt.cir.opt$par[3]

#******** rates
h.cir.opt.u=est_h_cir(kappa.cir.opt.u,sigma.cir.opt.u)
beta.cir.opt.u=est_beta_cir(h.cir.opt.u,kappa.cir.opt.u,ttm)
alpha.cir.opt.u=est_alpha.cir(h.cir.opt.u,kappa.cir.opt.u,theta.cir.opt.u,sigma.cir.opt.u,ttm)
spot.cir.opt.u=est_r_t(alpha.cir.opt.u,beta.cir.opt.u,ttm,r_t)

h.cir.opt=est_h_cir(kappa.cir.opt,sigma.cir.opt)
beta.cir.opt=est_beta_cir(h.cir.opt,kappa.cir.opt,ttm)
alpha.cir.opt=est_alpha.cir(h.cir.opt,kappa.cir.opt,theta.cir.opt,sigma.cir.opt,ttm)
spot.cir.opt=est_r_t(alpha.cir.opt,beta.cir.opt,ttm,r_t)

## instantaneous rates
beta_func_cir_u=function(ttm){(1-exp(-(-0.0000156*ttm)))/-0.0000156}

alpha_func_cir_u=function(ttm){exp((0.0478709-(0.000035175^2/(2*(-0.0000156^2))))*(beta_func_cir_u(ttm)-ttm)-
                                          (0.000035175^2/(4*-0.0000156))*beta_func_cir_u(ttm)^2)}

beta_func_cir=function(ttm){(1-exp(-(0.44977704*ttm)))/0.44977704}

alpha_func_cir=function(ttm){exp((0.04692863-(0.0001000^2/(2*(0.44977704^2))))*(beta_func_cir(ttm)-ttm)-
                                          (0.0001000^2/(4*0.44977704))*beta_func_cir(ttm)^2)}

## calculate the gradient
gb.u=grad(func=beta_func_cir_u,x=ttm,)
ga.u=grad(func=alpha_func_cir_u,x=ttm)
rslt.u=-ga.u + r_t*gb.u

gb=grad(func=beta_func_cir,x=ttm,)
ga=grad(func=alpha_func_cir,x=ttm)
rslt=-ga + r_t*gb

par(mfrow=c(1,2))
plot(r_t,type="l",lwd=2,main="constrained cir yields",ylim=range(cbind(spot.cir.opt,rslt)))
lines(spot.cir.opt,col="red",lwd=2)
lines(rslt,col="blue",lwd=2)
legend("bottomleft",legend=c("observed","spot - optim","instant - opim"),col=c("black","red","blue"),lwd=2)

plot(r_t,type="l",lwd=2,main="unconstrained cir yields",ylim=range(cbind(spot.cir.opt.u,rslt.u)))
lines(spot.cir.opt.u,col="red",lwd=2)
lines(rslt.u,col="blue",lwd=2)
legend("topright",legend=c("observed","spot - optim","instant - opim"),col=c("black","red","blue"),lwd=2)

#********** bond price
price_func_cir=function(ttm){alpha_func_cir(ttm) * exp(-beta_func_cir(ttm)*r_t)}
p_t_cir=price_func_cir(ttm)

#********** option price
bnd.beta.T.Tilde=beta_func_cir(opt.ttm)-beta_func_cir(bnd.ttm) # beta on 0.5 yr bond 2.5 yrs hence
bnd.alpha.T.Tilde=alpha_func_cir(opt.ttm)-alpha_func_cir(bnd.ttm)
bnd.p.t.Tilde=p_t_cir[which(ttm>3 & ttm<3.5)] #p_t_cir where 3 < ytm < 3.5
bnd.p.t.T=p_t_cir[which(ttm>2.5 & ttm<3)]

# bond option prices using cir
cir.rho=(2*h.cir.opt)/
     (sigma.cir.opt^2*(exp(h.cir.opt*opt.ttm)-1))

cir.psi=(kappa.cir.opt+h.cir.opt)/sigma.cir.opt^2

cir.mu=(log(bnd.alpha.T.Tilde/opt.strike))/
     bnd.beta.T.Tilde

cir.z=bnd.p.t.Tilde*pchisq(2*cir.mu*(cir.rho+cir.psi+bnd.beta.T.Tilde),
                           (4*kappa.cir.opt*theta.cir.opt)/sigma.cir.opt^2,
                           (2*cir.rho^2*bnd.r.t*exp(h.cir.opt*opt.ttm))/(cir.rho+cir.psi+bnd.beta.T.Tilde))-
     opt.strike*bnd.p.t.T*pchisq(2*cir.mu*(cir.rho+cir.psi),
                                 (4*kappa.cir.opt*theta.cir.opt)/sigma.cir.opt^2,
                                 (2*cir.rho^2*bnd.r.t*exp(h.cir.opt*opt.ttm))/(cir.rho+cir.psi))

as.numeric(cir.z*100)

```

